---
title: "Class 10 : Comparative structure analysis of Adenylate Kinase"
author: "Loreen A17059289"
format: html
---

# Comparative Analysis of ADK 

We will search the entire PDB for related structures using BLAST, fetch, align and superpose the identified structures, perform PCA and finally calculate the normal modes of each individual structure in order to probe for potential differences in structural flexibility.

Questions 10-12:

> Q10. Which of the packages above is found only on BioConductor and not CRAN?

msa

> Q11. Which of the above packages is not found on BioConductor or CRAN?:

bio3d-view

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?

TRUE


```{r}
library(bio3d)

aa <- get.seq("1ake_A")
aa
```

We can now run BLAST with this sequence:
```{r}
# Blast or hmmer search 
# b <- blast.pdb(aa)
```

```{r}
# hits <- plot(b)
```
> Q13. How many amino acids are in this sequence, i.e. how long is this sequence? 

214

Let's see what is in our `hits` object. Use pdb.id to access them. 

```{r}
# hits$pdb.id

hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

Now we can download all of these PDB structures:

```{r}
# Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
Now I want to align and superpose these structures which are all over the place. 

```{r}
# Align related PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```
Let's look at what `pdbs` looks like:

```{r}
# Alignment of all structures
pdbs
```

Now we have our aligned and superposed structures which we can perform all sorts of analysis on.
Let's do PCA:

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA on Adenylate kinase X-ray structures. Each dot represents one PDB structure.


We can cluster the structures by RMSD(or any other method):

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```
OPTIONAL:

We can make a movie - also called a trajectory of the major differences(i.e. structural displacemenets) of ADK.

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```



